A Constraint Solving Approach to Parikh Images of Regular Languages

### 2023-06-28 conclusion.tex
Conclusions and multi light touchup

%% For double-blind review submission, w/o CCS and ACM Reference (max submission space)
\documentclass[acmsmall,review,anonymous,screen]{acmart}\settopmatter{printfolios=true,printccs=true,printacmref=true}
%% For double-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
%% For single-blind review submission, w/o CCS and ACM Reference (max submission space)
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%% For single-blind review submission, w/ CCS and ACM Reference
%\documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true}
%% For final camera-ready submission, w/ required CCS and ACM Reference
%\documentclass[acmsmall]{acmart}\settopmatter{}

%% Journal information
%% Supplied to authors by publisher for camera-ready submission;
%% use defaults for review submission.
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2023}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission;
%% use 'none' for review submission.
\setcopyright{none}
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2018}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Note: Authors migrating a paper from PACMPL format to traditional
%% SIGPLAN proceedings format must update the '\documentclass' and
%% topmatter commands above; see 'acmart-sigplanproc-template.tex'.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Some recommended packages.
\usepackage{booktabs}   %% For formal tables:
                        %% http://ctan.org/pkg/booktabs
\usepackage{subcaption} %% For complex figures with subfigures/subcaptions
                        %% http://ctan.org/pkg/subcaption

\usepackage{amsmath,empheq,fancybox}
\usepackage{paralist}
\usepackage{url}
\usepackage{color}
\usepackage{textcomp,listings}
\usepackage{array}
\usepackage{mymacros}
\usepackage{microtype}
\usepackage{listings}
\usepackage{csquotes}
\usepackage{proof}
\usepackage[capitalise]{cleveref}
\usepackage{algorithm2e}      
\usepackage{multirow}
\usepackage{mathpartir}
\usepackage{amsthm}
\usepackage{numprint}
\usepackage{ebproof}
\usepackage[makeroom]{cancel}

\usepackage{mathtools} % Bonus

\usepackage{enumitem}

\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\newtheorem{example}{Example}[section]

\newlist{constraints}{enumerate}{3}
\setlist[constraints,1]{%
    label=\sffamily{(\roman*):},
    ref=\normalfont{Constraint (\roman*)},
    wide,itemsep=0pt,topsep=0pt
}    
\crefname{constraintsi}{}{}
\Crefname{constraintsi}{}{}

\lstset{
    columns=fullflexible,
    showspaces=false,
    showtabs=false,
    breaklines=true,
    showstringspaces=false,
    breakatwhitespace=true,
    escapeinside={(*@}{@*)},
    commentstyle=\color{greencomments},
    keywordstyle=\color{bluekeywords},
    stringstyle=\color{redstrings},
    numberstyle=\color{graynumbers},
    basicstyle=\ttfamily\small,
    framesep=12pt,
    xleftmargin=12pt,
    tabsize=4,
    captionpos=b
}

\newif\ifcomments%
\commentstrue%
\newif\ifoutline%
\outlinetrue{}

\newcommand{\contents}[1]{\ifoutline{\color{blue}
    \begin{itemize}
    #1
    \end{itemize}
  }\fi}

\allowdisplaybreaks[1]

\begin{document}

%% Title information
\title{A Constraint Solving Approach to Parikh Images of Regular Languages}
                                        %% when present, will be used in
                                        %% header instead of Full Title.
%% Author information
%% Contents and number of authors suppressed with 'anonymous'.
%% Each author should be introduced by \author, followed by
%% \authornote (optional), \orcid (optional), \affiliation, and
%% \email.
%% An author may have multiple affiliations and/or emails; repeat the
%% appropriate command.
%% Many elements are not rendered, but should be provided for metadata
%% extraction tools.

%% Author with single affiliation.
\author{First1 Last1}
\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position1}
  \department{Department1}              %% \department is recommended
  \institution{Institution1}            %% \institution is required
  \streetaddress{Street1 Address1}
  \city{City1}
  \state{State1}
  \postcode{Post-Code1}
  \country{Country1}                    %% \country is recommended
}
\email{first1.last1@inst1.edu}          %% \email is recommended

%% Author with two affiliations and emails.
\author{First2 Last2}
\authornote{with author2 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position2a}
  \department{Department2a}             %% \department is recommended
  \institution{Institution2a}           %% \institution is required
  \streetaddress{Street2a Address2a}
  \city{City2a}
  \state{State2a}
  \postcode{Post-Code2a}
  \country{Country2a}                   %% \country is recommended
}
\email{first2.last2@inst2a.com}         %% \email is recommended
\affiliation{
  \position{Position2b}
  \department{Department2b}             %% \department is recommended
  \institution{Institution2b}           %% \institution is required
  \streetaddress{Street3b Address2b}
  \city{City2b}
  \state{State2b}
  \postcode{Post-Code2b}
  \country{Country2b}                   %% \country is recommended
}
\email{first2.last2@inst2b.org}         %% \email is recommended

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
  A common problem in string constraint solvers is computing the Parikh image, a
   set of linear equations that describe all possible combinations of character
   counts in strings of a given language. Automata-based string solvers
   frequently need to compute the Parikh image of large products (intersections)
   of nondeterministic automata, which in many operations is both prohibitively
   slow and memory-intensive. We contribute a novel understanding of how Parikh
   maps can be tackled as a constraint solving problem to solve real-world
   constraints stemming from functions on regular languages, most notably the
   length constraint. Furthermore, we show how this formulation can be
   efficiently implemented as a calculus, \Calculus{}, in an automated theorem
   prover supporting Presburger logic.  We implement \Calculus{} in a tool
   called \Catra{}, and evaluate it on constraints produced by the
   \OstrichPlus{} string constraint solver when solving Parikh automata
   intersection problems produced when solving standard string constraint
   benchmarks involving string length constraints. We show that our solution
   strictly outperforms the standard approach described
   in~\citeauthor{generate-parikh-image} as well as the over-approximating
   method recently described by~\citeauthor{approximate-parikh} by a wide
   marigin, and for realistic timeouts for constraint solving also the~\Nuxmv{}
   model checker.
\end{abstract}

%% 2012 ACM Computing Classification System (CSS) concepts
%% Generate at 'http://dl.acm.org/ccs/ccs.cfm'.
\begin{CCSXML}
  <ccs2012>
  <concept>
  <concept_id>10003752.10003766.10003773.10003775</concept_id>
  <concept_desc>Theory of computation~Quantitative automata</concept_desc>
  <concept_significance>500</concept_significance>
  </concept>
  <concept>
  <concept_id>10003752.10003766.10003776</concept_id>
  <concept_desc>Theory of computation~Regular languages</concept_desc>
  <concept_significance>300</concept_significance>
  </concept>
  <concept>
  <concept_id>10003752.10003790.10003794</concept_id>
  <concept_desc>Theory of computation~Automated reasoning</concept_desc>
  <concept_significance>500</concept_significance>
  </concept>
  <concept>
  <concept_id>10003752.10010124.10010138.10010142</concept_id>
  <concept_desc>Theory of computation~Program verification</concept_desc>
  <concept_significance>300</concept_significance>
  </concept>
  <concept>
  <concept_id>10003752.10003790.10011192</concept_id>
  <concept_desc>Theory of computation~Verification by model checking</concept_desc>
  <concept_significance>300</concept_significance>
  </concept>
  </ccs2012>  
\end{CCSXML}

\ccsdesc[500]{Theory of computation~Quantitative automata}
\ccsdesc[300]{Theory of computation~Regular languages}
\ccsdesc[500]{Theory of computation~Automated reasoning}
\ccsdesc[300]{Theory of computation~Program verification}
\ccsdesc[300]{Theory of computation~Verification by model checking}
%% End of generated code

%% Keywords
%% comma separated list
\keywords{Parikh images, string solvers, model checking}  %% \keywords are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\section{Introduction}\label{sec:introduction}
\input{introduction.tex}

\section{An Intuition for Our Approach}\label{sec:intuition}
\input{example.tex}

\section{Preliminaries}\label{sec:preliminaries}
\input{preliminaries.tex}

\section{Projections on Parikh Images}\label{sec:generalised}
\input{generalised.tex}

\section{A Calculus for Projections on Parikh Images}\label{sec:calculus}
\input{calculus.tex}

\section{Parikh Images from Products of Automata}\label{sec:multiple}
\input{multiple.tex}

\section{Extensions}\label{sec:extensions}
\input{extensions.tex}

\section{Implementation}\label{sec:implementation}
\input{implementation.tex}

\section{Evaluation}\label{sec:experiments}
\input{evaluation.tex}

\section{Conclusion}

@@

Within \Catra{}, \Calculus{} shows astonishing performance in terms of memory
usage and solve-time compared to the baseline approach laid out in
\cite{generate-parikh-image} when implemented on the same underlying automated
theorem prover (\Princess{}, \cite{princess}). It is also competitive with the
\Nuxmv{} model checker \cite{nuxmv}, outperforming it on unsatisfiable instances
and generally outperforming it for timeouts under 30 seconds with its advantage
increasing drastically for even shorter timeouts. 30 seconds would generally be
considered a long timeout for our intended use as supporting infrastructure to a
string constraint solver.

@@

%% Acknowledgments
% \begin{acks}                            %% acks environment is optional
%                                         %% contents suppressed with 'anonymous'
%   %% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
%   %% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
%   %% acknowledge financial support and will be used by metadata
%   %% extraction tools.
%   This material is based upon work supported by the
%   \grantsponsor{GS100000001}{National Science
%     Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
%   No.~\grantnum{GS100000001}{nnnnnnn} and Grant
%   No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
%   conclusions or recommendations expressed in this material are those
%   of the author and do not necessarily reflect the views of the
%   National Science Foundation.
% \end{acks}

%% Bibliography
\bibliography{bibliography}

%% Appendix
\appendix
\section{Appendix}
\input{appendix.tex}
\end{document}

>>>>

@@

Within \Catra{}, \Calculus{} shows astonishing performance in terms of
solve-time compared to the baseline approach laid out in
\cite{generate-parikh-image} when implemented on the same underlying automated
theorem prover (\Princess{}, \cite{princess}). It is also competitive with the
\Nuxmv{} model checker \cite{nuxmv}, outperforming it on unsatisfiable instances
and generally outperforming it for timeouts under 30 seconds with its advantage
increasing drastically for even shorter timeouts. 30 seconds would generally be
considered a long timeout for our intended use as supporting infrastructure to a
string constraint solver.

@@

### 2023-06-28 main.tex
Conclusions and multi light touchup

\section{Conclusion}

In this paper we have introduced a calculus to compute commutating operations on
intersections of regular languages that we call \Calculus{}. We have evaluated
it on \NrBenchmarks{} Parikh automata intersection problems generated by the
\OstrichPlus{} string solver \cite{ostrich-plus} solving the PyEx benchmark
suite \cite{pyex} using our Parikh automata solver \Catra{}.

Within \Catra{}, \Calculus{} shows astonishing performance in terms of memory
usage and solve-time compared to the baseline approach laid out in
\cite{generate-parikh-image} when implemented on the same underlying automated
theorem prover (\Princess{}, \cite{princess}). It is also competitive with the
\Nuxmv{} model checker \cite{nuxmv}, outperforming it on unsatisfiable instances
and generally outperforming it for timeouts under 30 seconds with its advantage
increasing drastically for even shorter timeouts. 30 seconds would generally be
considered a long timeout for our intended use as supporting infrastructure to a
string constraint solver.

Future investigations involve two tracks. The first one is integration into
existing string solvers (wich \Ostrich{} being a particularly promising
candidate due to its shared use of \Princess{}), and further adaptation to that
use case. Closer inspection of the instances where we currently time out should
be useful to further improve our heuristics.

The second track for future improvements is the extension into other problem
domains, including other logics, model checking problems, as well as to more
powerful automata such as transducers. In principle, we are also already able to
express stronger constraints than Parikh automata, due to our use of a full
automated theorem prover which allows adding arbitrary constraints in addition
to the expected Presburger formulae.

>>>>

\section{Conclusion}
\input{conclusion.tex}
